---
title: "PS270 Final Project"
author: "Zain Ashraf"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---
## Introduction

There has been much discourse about the desirability of walkable and bikeable cities with good transit. The appeal of such cities transcends the fact that an individual does not have to own a car and manage the expenses associated with it. Cities where walking and biking are feasible transportation options can promote more healthy, active lifestyles than those where residents are resigned to spending hours each week driving to their destinations. Furthermore, cities with less cars have less pollution, and many individuals prefer the dense urban planning that often comes with more connected cities. While cities that are well connected are very desirable, many individuals still opt to live in car-dependent areas with weak walkability, bikeability, and public transportation options. While each person’s reasoning may be different, it is the trend that cities with more Urban Connectivity are more expensive. While this is an idea supported by cost of living trends, I decided to further investigate if Urban Connectivity was the reason for these Cost of Living Increases. I hypothesize that the higher Urban Connectivity a city has, the higher Cost of Living it will have.

Below, I have loaded packages that may include necessary functions for my analysis.

```{r}
library(tidyverse)
library(infer)
library(broom)
library(modelsummary)
```

## Data
I merged two datasets to measure the relationship between the cost of living and urban connectivity of various cities in the United States. My independent variable is Urban Connectivity and my dependent variable is Cost of Living. This study follows a cross-sectional design as these datasets each contain data from one point in time.

I found my first dataset, titled “U.S. Cities Urban Connectivity,” on Kaggle (Vivian Ellis, 2024. https://doi.org/10.34740/KAGGLE/DSV/8715704). It contains data on various United States cities’ walkability, public transportation availability, and biking infrastructure. These are considered the factors that determine a city’s urban connectivity. These factors were measured through expert coding by walkscore.com. To understand the urban connectivity of a city as a whole, I will sum these factors to create a broad “Urban Connectivity Score”. This score will be my independent variable that I will measure against the Cost of Living Index.

I sourced my second dataset, titled “AdvisorSmith Cost of Living Index” at https://advisorsmith.com/data/coli. This dataset contains the Cost of Living data for my study. AdvisorSmith created an index that describes a city’s cost of living in relation to the average cost of living in the United States. A score of 100 represents the average cost of living. A score of 125 represents a cost of living 25% higher than average. A score of 75 represents a cost of living 25% less than average. Thus, this variable is measured through expert coding. Cost of Living will be my dependent variable.


Below is the code I used to load in the two datasets and merge them.
Setting working directory.

```{r}
setwd("C:/Users/Zain/Desktop/ps270final")
```

Loading in cost of living (COL) dataset.

```{r}
coli <-
read_csv("data/advisorsmith_cost_of_living_index.csv")
```

Loading in urban connectivity dataset.

```{r}
connectivity <-
read_csv("data/urban_connectivity.csv")
```

Merging datasets to have one with both COL and Urban Connectivity.

```{r}
cities <- coli |>
  inner_join(connectivity) |>
  select(City, State, `Cost of Living Index`, Place_name, `Walk Score`, `Transit Score`, `Bike Score`, Population_2022_Census)
```

A glimpse of the dataset.

```{r}
glimpse(cities)
```

To visualize the spread of Cost of Living represented in the data, I have generated a boxplot.

```{r}
cities |>
  summarize(avg_col = mean(`Cost of Living Index`)) 
```

```{r}
col_box <- cities |>
  ggplot(aes(x = `Cost of Living Index`)) +
  geom_boxplot() +
  geom_vline(xintercept = 103.6958, linetype = "dashed", color = "dodgerblue") +
  labs(title = "Spread of Costs of Living (COL)",
       caption = "Blue Dotted Line: Average COL")
col_box
```

The sample has an median Cost of Living close the national average Cost of Living. The average Cost of living of the dataset is 3% higher than the national average. 

## Results
To compare  Urban Connectivity to Cost of Living, rather than a component of Urban Connectivity, I need to sum the each city's walk score, transit score, and bike score. Doing so will provide me a single value for each city that measures its overall connectivity.

Creating Urban Connectivity variable ("Total Connectivity")

```{r}
cities <- cities |>
  mutate("Total Connectivity" = `Walk Score` + `Transit Score` + `Bike Score`) |>
  filter(`Total Connectivity` > 0)
glimpse(cities)
```

After removing the cities with missing Transit Scores, causing incomplete Total Connectivity scores, I am left with 66 cities from which I will conduct my analysis.

First, I will create a scatterplot visualizing the correlation between Total Connectivity and Cost of Living.

```{r}
cities |>
  ggplot(aes(y = `Cost of Living Index`,
             x = `Total Connectivity`,
             color = State)) +
  geom_point() 
```
The scatterplot indicates a weak, but positive correlation between Total Connectivity and Cost of Living Index. This means that as the urban connectivity (i.e., walkability, bikeability, and transit options) of a city increases, so does its Cost of Living compared to the average Cost of Living in the United States. However, the correlation is relatively weak, so I will conduct a regression which includes a possible confounder, population size. Population size may influence the Cost of Living of a city because a higher population leads to increased demand for housing, food, and healthcare, all of which are factored into AdvisorSmith's calculation for cost of living.

Here is my regression, with Cost of Living as the outcome variable and Total Connectivity as the main predictor variable, with City Population (according to the 2022 census) as a potential confounding variable. 

```{r}
options(scipen = 999) #disables scientific notation
cities_reg <- lm(data = cities, `Cost of Living Index` ~ `Total Connectivity` + Population_2022_Census)
```

To find the p-value and represent the regression in a nice table, I will use model summary.

```{r}
modelsummary(cities_reg,
             statistic = c("s.e. = {std.error}",
                           "p = {p.value}"),
             gof_map = c("nobs", "r.squared", "adj.r.squared"))
```
The effect of Urban Connectivity on a city's Cost of Living is significant. Because the p-value is extremely close to 0, we can reject a null hypothesis that a city's Urban Connectivity has no effect on its Cost of Living at any alpha level. Therefore, we can interpret the effect as "for every 1-point increase in a city's Total Urban Connectivty score, its Cost of Living becomes 0.22% higher compared to the national Cost of Living average. The population size does not represent a confounding variable to this effect.

## Conclusion

# Results

The regression indicates a significant effect of Urban Connectivity on Cost of Living. While the correlation is relatively weak, being only .228, my hypothesis that cities with higher levels of Urban Connectivity are more expensive is supported. The effect size (R2) indicates that 27.8% of the variance in cities Cost of Living can be explained by their Urban Connectivity. 

# Limitations 

The connectivity data is relatively recent. The cost of living data is not, however, being from 2021. That said, it is a relative measure so relationships between cities should probably be the same. The population data is from 2022. Essentially, this study is limited by the close, yet different years in which the data was collected. Another limitation is the sample size. 66 cities were ultimately used for the analysis, and it is unknown if these cities constitute a representative spread of cities across the United States. However, they do establish a trend. 

# Areas of improvement

I think the study would benefit from conducting more analyses of potential confounding variables, such as location, whether state or regionally. Additionally, as mentioned before, the study would benefit from having more cities represented in the dataset.

